# Websites using pkgdown, bookdown, and blogdown

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = FALSE)
```

In order to serve html pages through GitHub pages the following needs to be done to create an emtpy gh-pages branch to be used to deploy the html pages created by either `pkgdown`, `bookdown`, or `blogdown`

Derived from: https://pkgdown.r-lib.org/reference/deploy_site_github.html

First we need to create an empty gh-pages branch. In bash:
  
  ```
git checkout --orphan gh-pages
git rm -rf .
git commit --allow-empty -m 'Initial gh-pages commit'
git push origin gh-pages
git checkout master
```

## Action to deploy a pkgdown site

Now we can add in pkgdown

```{r}
usethis::use_pkgdown()
```

Then we want to add in the github action for pkgdown to automatically add documentation
```{r}
usethis:::use_github_action(url = "https://raw.githubusercontent.com/r-lib/actions/master/examples/pkgdown.yaml")
```

N.B. The usethis function is not currently exported into the NAMESPACE of `usethis` but is documented.

## Action to deploy a bookdown site

## Action to deploy blogdown site